Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-09-06 pnas/Towards.tex
small changes

The signal to noise ratio (SNR) of the projections in the micrographs is a function of two dominating factors. On the one hand, the SNR is a function of the electron dose. To keep radiation damage within acceptable bounds, the dose must be kept low, which leads to high noise levels. On the other hand, the SNR is a function of the molecule size. The smaller the molecules, the fewer detected electrons carry information about them. \TODO{This isn't quite accurate. Amit: do you agree with the last sentence? maybe we should say it is a function of the contrast?}

@@

Setting $W = L$ (as is a priori desired) is problematic because the above optimization problems appears to have numerous poor local optimizers. Similar phenomenon was recently observed in a related problem~\cite{zhang2018structured}.
Thus, we first run the optimization with $W = 2L-1$. This problem appears to have few poor local optima, perhaps because the additional degrees of freedom allow for more escape directions. Since we hope the signals estimated this way correspond to the true signals zero-padded to length $W$, we extract from each one a subsignal of length $L$ that has largest $\ell_2$-norm. This estimator is then used as initial iterate for~\eqref{eq:optim1D}, this time with $W = L$. We find that this procedure is reliable for a wide range of experimental parameters. To solve~\eqref{eq:optim1D}, we run the trust-region method implemented in Manopt~\cite{manopt}, which allows to treat the positivity constraints on coefficients $\hat \gamma$. Notice that the cost function is a polynomial in the variables, so that it is straightforward to compute it and its derivatives.

@@

The micrographs for the experiment presented in Figure~\ref{fig:cryo_detection} were generated as follows. We sample rotation matrices from SO(3) uniformly at random using the QR-based algorithm described in~\cite{stewart1980efficient}, and generate the projection of the volume corresponding to that rotation matrix using \verb|cryo_project| in ASPIRE\footnote{\url{http://spr.math.princeton.edu}}. %\cite{aspire}.
 We keep track of the indices at which the upper left corner of a projection can be placed without violating the separation condition, so all projections are separated by at least 30 pixels in each dimension. We continue adding projections to the micrograph until no more projections can be added without violating the separation condition. Then, the LS problem~\eqref{eq:LS_cryo} was solved assuming the spherical harmonic cutoff for the volume is $L=0$, which is sufficient to recover a significant $\widehat\gamma$ in the presence of projections in the micrograph.

@@

We start by considering the mean of the data:
\begin{equation*}
a_y^1 = \frac{1}{N}\sum_{i=0}^{N-1} y[i] =
\frac{1}{N/L}\sum_{j=0}^{M-1}\frac{1}{L}\sum_{i=0}^{L-1}x[i] +    
\underbrace{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]}_{\text{noise term}}
\xrightarrow{a.s.}\gamma a_x^1,
\end{equation*}
%
where the noise term converges to zero almost surely (a.s.) by the strong law of large numbers.

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
\\ & + \underbrace{\frac{1}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] (\varepsilon[s_j + i + \ell] + \varepsilon[s_j + i - \ell])}_{\text{cross-term}}. 
%
\end{align*}

The cross-term is linear in the noise, and is easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. We break the signal term into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{align} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] & = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\nonumber\\&\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{align}
%

We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation*}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, hence converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
%
\end{equation*}

@@

The proof is based on  a variant of the Neyman--Pearson Lemma to derive the best (deterministic) estimator $\hat{\eta}$. We take any estimator $\hat{\eta}$, and we let $S$ be the set of $X$'s where $\hat{\eta} = 1$. We write $\P_i$ to mean the probability conditional on the event $\eta = i$; that is, $\P_i[A] = \P[A | \eta = i]$. Then, the probability that $\hat{\eta}$ fails is:

@@

The best estimator of $\eta$ based on $X$ minimizes the failure probability; hence, it minimizes the integral in~\eqref{eq:neyman-pearson} through an appropriate choice of the set $S$. This is achieved as follows:

@@

S & = \left\{x : qf_0(x) \leq (1-q)f_1(x) \right\}.

@@

To complete the proof, we need to show that the two quadratic equations~\eqref{eq:quad1} and~\eqref{eq:quad2} are independent. To this end, it is enough to show that the ratios between coefficients differ. 
From~\eqref{eq:quad1} and~\eqref{eq:E1}, we have:
\begin{equation*}
\begin{split}
\frac{\mathcal{B}}{\mathcal{C}} &= \frac{E_1 - (2L+1)(a_y^1)^3 - a_y^1E_3}{(a_y^1)^3} \\&= \frac{a_y^2[0] - (2L+1)(a_y^1)^2 - E_3}{(a_y^1)^2}.
\end{split}
\end{equation*}
In addition, using~\eqref{eq:E3},
\begin{equation*}
\frac{\mathcal{E}}{\mathcal{F}} = \frac{(3-6L)(a_y^1)^3}{LE_4} = 3 - 6L . 
\end{equation*}
For contradiction, suppose that the quadratics are dependent. Then, $\frac{\mathcal{B}}{\mathcal{C}} =\frac{\mathcal{E}}{\mathcal{F}} $, that is, 	
\begin{equation*}
a_y^2[0] - (2L+1)(a_y^1)^2 - E_3 = (a_y^1)^2(3-6L).
\end{equation*}
Rewriting the identity in terms of $x$ and dividing by $\beta$ we get:
\begin{equation} \label{eq:cond}
4(L-1)\beta (\1^\top x)^2  - (\1^\top x)^2 + \|x\|^2 = 0.
\end{equation}	
% Nicolas, Sep 4 2018: everything checked and correct until here.
For generic $x$,  this polynomial equation is not satisfied so that the quadratic equations are independent. 
Furthermore, from the inequality $L\|x\|^2 \ge (\1^\top x)^2$ it follows immediately that the equations must be independent so long as
\begin{equation*}
\beta > \frac{1}{4L}.
\end{equation*}

@@

We begin by calculating the relation between the $p$th autocorrelation of the clean micrograph and the  averaged autocorrelation of the projections.
Let us denote the clean micrograph by $\widetilde \II = \II-\varepsilon$, where $\II$ and $\varepsilon$ are given in~\eqref{eq:micrograph_model}.     
Denote by $\mc S_t$ the support of the $t$th particle. 
Then, we have
\begin{equation} \label{eq:general_moment}
\begin{split}
	a^p_{\widetilde\II}[\Bell_1,\ldots, \Bell_{p-1}] &= \frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\widetilde\II[\mb i + \Bell_1]\cdots \widetilde\II[\mb i + \Bell_{p-1}]\\
&= \frac{1}{N^2}\sum_{t=1}^M\sum_{\mb i\in\mc S_t}\widetilde\II[\mb i]\widetilde\II[\mb i + \Bell_1]\cdots \widetilde\II[\mb i + \Bell_{p-1}] \\ %+ \frac{1}{N^2}\sum_{\mb i\in Z}\widetilde\II[\Bell]\widetilde\II[\mb i + \Bell]\cdots \widetilde\II(\mb i + \Bell_{p-1})\\
&= \frac{MP^2}{N^2}\cdot\frac{1}{M}\sum_{t=1}^M\frac{1}{P^2}\sum_{i,j=0}^{P-1} I_{\omega_t}[\mb i]I_{\omega_t}[\mb i+\Bell_1]\cdots I_{\omega_t}[\mb i + \Bell_{p-1}]\\
&= \frac{MP^2}{N^2}\frac{1}{M}\sum_{t=1}^Ma^p_{I_{\omega_t}}[\Bell_1,\ldots, \Bell_{p-1}]\\
&\to \gamma\langle a^p_{I_{\omega}}[\Bell_1,\ldots, \Bell_{p-1}]\rangle_{\omega},
\end{split}
\end{equation}
where the average is taken over $\omega$ with respect to the distribution of viewing directions. Here, we assume it to be uniform.

@@

For the third moments, we get 8 terms:
\[\begin{aligned} 
&a^3_\II[\Bell_1, \Bell_2]  \\&
 = \underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\widetilde\II[\mb i+\Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(1)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\varepsilon[\mb i+\Bell_1]\varepsilon[\mb i + \Bell_2]}_{(2)}\\ 
&+ \underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\varepsilon[\mb i + \Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(3)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\widetilde\II[\mb i + \Bell_1]\varepsilon[\mb i + \Bell_2]}_{(4)}\\
&+ \underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\widetilde\II[\mb i + \Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(5)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\varepsilon[\mb i + \Bell_1]\varepsilon[\mb i + \Bell_2]}_{(6)}\\
&+ \underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\varepsilon[\mb i + \Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(7)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\widetilde\II[\mb i + \Bell_1]\varepsilon[\mb i + \Bell_2]}_{(8)}.
\end{aligned}\]
We address these terms one by one:
\begin{itemize}
	\item  Term (1) is treated by~\eqref{eq:general_moment} for $p=3$;
	\item Term (2) is the third-order autocorrelation of  pure noise  which vanishes in the limit; 
	\item Terms (3)-(5) depend linearly on the noise and hence
	vanish in the limit;
	\item For term (6), if $\Bell_1\neq \Bell_2$ the term
	vanishes in the limit. If $\Bell_1 = \Bell_2$ then
	\[ \begin{aligned}
	\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\varepsilon[\mb i + \Bell]^2
	&= \frac{MP^2}{N^2}\cdot\frac{1}{MP^2}\sum_{t=1}^{M}\sum_{\mb i\in\mc S_t}I_{\omega_t}[\mb i]\varepsilon[\mb i + \Bell]^2 
	\\& \to \gamma\sigma^2 \langle a^1_{I_{\omega}}\rangle_{\omega}, %\gamma\mathbb{E}_{\omega, \varepsilon}[\varepsilon^2I_{\omga}]\\
	%&= \gamma\sigma^2\mathbb{E}[m_1[I_{\omega}]]\\ 
	%&= \gamma\sigma^2Lm_1[\phi],
	\end{aligned}\]
	where $\langle a^1_{I_{\omega}}\rangle_{\omega}$ is the mean of the volume. 
	\item  Terms (7) and (8) contribute $\delta$ functions similar to (6).
\end{itemize}
Thus, we conclude that
\begin{equation} \label{eq:third_order_ac_micro_volume}
a^3_\II[\Bell_1, \Bell_2] \to \gamma\langle
a^3_{I_{\omega}}[\Bell_1, \Bell_2]\rangle_{\omega} +
\gamma\sigma^2\langle a^1_{I_{\omega}}\rangle_{\omega}\Big(\delta[\Bell_1 - \Bell_2] +
\delta[\Bell_1] + \delta[\Bell_2]\Big),
\end{equation}
where the second term is the bias $b_3[\Bell_1,\Bell_2]$.
Note that $\gamma\langle a^1_{I_{\omega}}\rangle_{\omega}$ is approximately the mean of the micrograph since  $a^1_\II \approx a^1_{\tilde\II} \approx \gamma\langle a^1_{I_{\omega}}\rangle_{\omega}$. Therefore, we do not need prior knowledge of $\gamma$ to effectively
debias the third-order autocorrelation.

@@

We represent our autocorrelations using Prolate Spheroidal Wave Functions (PSWFs) $\{\psi_{k,q}\}$ where $k \geq 0, q \geq 1$ are integers~\cite{slepian1964pswfIV}. As we demonstrate below, this makes it easier to account for the fact that all in-plane rotations of the micrographs are equally likely observations. This is only a concern for the second and third order autocorrelations. Below, we start with $p=2$.
The PSWFs are eigenfunctions of the truncated Fourier transform and are given in polar coordinates by \footnote{A different normalization is used in \cite{landa2017steerable}.}
\begin{align}
\psi_{k,q}(r,\varphi) & = \left\{\begin{array}{ll} \frac{1}{\sqrt{8\pi}}\alpha_{k,q}R_{k,q}(r)e^{\iota k\varphi}, & r\leq 1,\\ 0, & r>1,\end{array}\right. \label{eq:prolatesdef}
\end{align}
where the range of $k,q$ is determined by Eq.~(8) in \cite{landa2017steerable}, the ${R_{k,q}}$ are a family of real, one-dimensional functions and the ${\alpha_{k,q}}$ are  scaling factors which will be defined in the next section.
The PSWFs are orthogonal on the unit disk.

@@

We now turn out attention to the third order autocorrelation. Following the same lines, we define the coefficients:
\[\begin{aligned} \mathfrak{a}^3[k_1,q_1;k_2,q_2] &= \sum_{\Bell_1, \Bell_2} a^3_\II[\Bell_1,\Bell_2]\overline{\psi_{k_1,q_1}[\Bell_1]}\psi_{k_2,q_2}[\Bell_2]\\
&= \frac{1}{N^2}\sum_{\mb i}\II[\mb i]\left(\sum_{\Bell_1}\II[\mb i+\Bell_1]\overline{\psi_{k_1,q_1}[\Bell_1]}\right)\\&\times \left(\sum_{\Bell_2}\II[\mb i+\Bell_2]\psi_{k_2,q_2}[\Bell_2]\right)\\
&= \frac{1}{N^2}\sum_{\mb i}\II[\mb i] a_{k_1,q_1}[\mb i]\overline{a_{k_2,q_2}[\mb i]},
\end{aligned}\]
where the patch expansion coefficients $a_{k,q}$ are as defined in~\eqref{eq:patchcoefficients}.
The coefficients $\mathfrak{a}^3[k_1,q_1;k_2,q_2]$ are related to the third-order autocorrelation via the approximate identity:
\[ a^3_\II[\Bell_1, \Bell_2] \approx \sum_{\substack{k_1,q_1\\ k_2,q_2}}\mathfrak{a}^3[k_1,q_1;k_2,q_2]\psi_{k_1,q_1}[\Bell_1]\overline{\psi_{k_2,q_2}[\Bell_2]}.\]
Averaging over all rotations of $\II$ and its reflection, we obtain
\begin{align} \label{eq:third_order_ac_pswf}
&\mathfrak{a}^3[k_1,q_1;k_2,q_2] = \frac{1}{N^2}\sum_{\mb i}\II[\mb i]\\&\times \left(\frac{1}{4\pi}\int_0^{2\pi}\left(a_{k_1,q_1}[\mb i]\overline{a_{k_2,q_2}[\mb i]} + \overline{a_{k_1,q_1}[\mb i]}a_{k_2,q_2}[\mb i]\right)e^{-\iota(k_1-k_2)\alpha}d\alpha\right), \nonumber\\
&= \delta[k_1 - k_2]\frac{1}{N^2}\sum_{\mb i}\II[\mb i]\Re\{a_{k_1,q_1}[\mb i]\overline{a_{k_2,q_2}[\mb i]}\}. \nonumber 
\end{align}
Thus, similarly to the second-order autocorrelation, averaging over all in-plane rotations reveals that the 4-D third-order autocorrelation truly only carries information along three dimensions.

>>>>

The signal to noise ratio (SNR) of the projections in the micrographs is a function of two dominating factors. On the one hand, the SNR is a function of the electron dose. To keep radiation damage within acceptable bounds, the dose must be kept low, which leads to high noise levels. On the other hand, the SNR is a function of the molecule size. The smaller the molecules, the fewer detected electrons carry information about them. \TODO{Amit: do you agree with the last sentence? maybe we should say it is a function of the contrast?}

@@

Setting $W = L$ (as is a priori desired) is problematic because the above optimization problems appears to have numerous poor local optimizers. A similar phenomenon was recently observed in a related problem~\cite{zhang2018structured}.
Thus, we first run the optimization with $W = 2L-1$. This problem appears to have few poor local optima, perhaps because the additional degrees of freedom allow for more escape directions. Since we hope the signals estimated this way correspond to the true signals zero-padded to length $W$, we extract from each one a subsignal of length $L$ that has largest $\ell_2$-norm. This estimator is then used as initial iterate for~\eqref{eq:optim1D}, this time with $W = L$. We find that this procedure is reliable for a wide range of experimental parameters. To solve~\eqref{eq:optim1D}, we run the trust-region method implemented in Manopt~\cite{manopt}, which allows to treat the positivity constraints on coefficients $\hat \gamma$. Notice that the cost function is a polynomial in the variables, so that it is straightforward to compute it and its derivatives.

@@

The micrographs for the experiment presented in Figure~\ref{fig:cryo_detection} were generated as follows. We sample rotation matrices from SO(3) uniformly at random using the QR-based algorithm described in~\cite{stewart1980efficient}, and generate the projection of the volume corresponding to that rotation matrix using \verb|cryo_project| in ASPIRE\footnote{\url{http://spr.math.princeton.edu}}. %\cite{aspire}.
 We keep track of the indices at which the upper left corner of a projection can be placed without violating the separation condition, so all projections are separated by at least 30 pixels in each dimension. We continue adding projections to the micrograph until no more projections can be added without violating the separation condition. \TODO{Nico: is this really how the procedure works? Or is it just that after a few failed attempts, we abandon? Also, the text doesn't say if the projections are placed at random, but I'm assuming they are ?} Then, the LS problem~\eqref{eq:LS_cryo} was solved assuming the spherical harmonic cutoff for the volume is $L=0$, which is sufficient to recover a significant $\widehat\gamma$ in the presence of projections in the micrograph.

@@

We start by considering the mean of the data:
\begin{equation*}
a_y^1 = \frac{1}{N}\sum_{i=0}^{N-1} y[i] =
\frac{1}{N/L}\sum_{j=0}^{M-1}\frac{1}{L}\sum_{i=0}^{L-1}x[i] +    
\underbrace{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]}_{\text{noise term}}
\xrightarrow{a.s.}\gamma a_x^1,
\end{equation*}
%
where the noise term converges to zero almost surely (a.s.)\ by the strong law of large numbers.

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
\\ & + \underbrace{\frac{1}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] (\varepsilon[s_j + i + \ell] + \varepsilon[s_j + i - \ell])}_{\text{cross-term}}. 
%
\end{align*}
The cross-term is linear in the noise, and is easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. We break the signal term into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{align} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] & = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\nonumber\\&\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{align}
%
We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation*}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, hence converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
%
\end{equation*}

@@

The proof is based on  a variant of the Neyman--Pearson Lemma to derive the best (deterministic) estimator $\hat{\eta}$. Take any estimator $\hat{\eta}$; it is characterized by $S$: the set of $X$'s where $\hat{\eta} = 1$, where $X$ is a random vector as defined above Proposition~\ref{prop:two_gauss}. We write $\P_i$ to mean the probability conditional on the event $\eta = i$; that is, $\P_i[A] = \P[A | \eta = i]$. Then, the probability that $\hat{\eta}$ fails is:

@@

The best estimator of $\eta$ based on $X$ minimizes the failure probability; hence, it minimizes the integral in~\eqref{eq:neyman-pearson} through an appropriate choice of the set $S$. This is achieved by picking all $x$'s such that the integrand is nonpositive:

@@

S & = \left\{x : qf_0(x) - (1-q)f_1(x) \leq 0 \right\}.

@@

To complete the proof, we need to show that the two quadratic equations~\eqref{eq:quad1} and~\eqref{eq:quad2} are independent. To this end, it is enough to show that the ratios between coefficients differ. 
From~\eqref{eq:quad1} and~\eqref{eq:E1}, we have:
\begin{equation*}
\begin{split}
\frac{\mathcal{B}}{\mathcal{C}} &= \frac{E_1 - (2L+1)(a_y^1)^3 - a_y^1E_3}{(a_y^1)^3} \\&= \frac{a_y^2[0] - (2L+1)(a_y^1)^2 - E_3}{(a_y^1)^2}.
\end{split}
\end{equation*}
In addition, using~\eqref{eq:E3},
\begin{equation*}
\frac{\mathcal{E}}{\mathcal{F}} = \frac{(3-6L)(a_y^1)^3}{LE_4} = 3 - 6L . 
\end{equation*}
For contradiction, suppose that the quadratics are dependent. Then, $\frac{\mathcal{B}}{\mathcal{C}} =\frac{\mathcal{E}}{\mathcal{F}} $, that is, 	
\begin{equation*}
a_y^2[0] - (2L+1)(a_y^1)^2 - E_3 = (a_y^1)^2(3-6L).
\end{equation*}
Rewriting the identity in terms of $x$ and dividing by $\beta$ we get:
\begin{equation} \label{eq:cond}
4(L-1)\beta (\1^\top x)^2  - (\1^\top x)^2 + \|x\|^2 = 0.
\end{equation}	
For generic $x$,  this polynomial equation is not satisfied so that the quadratic equations are independent. 
Furthermore, from the inequality $L\|x\|^2 \ge (\1^\top x)^2$ it follows immediately that the equations must be independent so long as
\begin{equation*}
\beta > \frac{1}{4L}.
\end{equation*}
% Nicolas, Sep 4 2018: everything checked and correct until here.

@@

We begin by calculating the relation between the $p$th autocorrelation of the clean micrograph and the  averaged autocorrelation of the projections.
Let us denote the clean micrograph by $\widetilde \II = \II-\varepsilon$, where $\II$ and $\varepsilon$ are given in~\eqref{eq:micrograph_model}.     
Denote by $\mc S_t$ the support of the $t$th particle. 
Then, we have
\begin{align}
	& a^p_{\widetilde\II}[\Bell_1,\ldots, \Bell_{p-1}]  \nonumber\\
	&= \frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\widetilde\II[\mb i + \Bell_1]\cdots \widetilde\II[\mb i + \Bell_{p-1}]\nonumber\\
&= \frac{1}{N^2}\sum_{t=1}^M\sum_{\mb i\in\mc S_t}\widetilde\II[\mb i]\widetilde\II[\mb i + \Bell_1]\cdots \widetilde\II[\mb i + \Bell_{p-1}] \nonumber\\ %+ \frac{1}{N^2}\sum_{\mb i\in Z}\widetilde\II[\Bell]\widetilde\II[\mb i + \Bell]\cdots \widetilde\II(\mb i + \Bell_{p-1})\\
&= \frac{MP^2}{N^2}\cdot\frac{1}{M}\sum_{t=1}^M\frac{1}{P^2}\sum_{i,j=0}^{P-1} I_{\omega_t}[\mb i]I_{\omega_t}[\mb i+\Bell_1]\cdots I_{\omega_t}[\mb i + \Bell_{p-1}]\nonumber\\
&= \frac{MP^2}{N^2}\frac{1}{M}\sum_{t=1}^Ma^p_{I_{\omega_t}}[\Bell_1,\ldots, \Bell_{p-1}]\nonumber\\
&\to \gamma\langle a^p_{I_{\omega}}[\Bell_1,\ldots, \Bell_{p-1}]\rangle_{\omega},
\label{eq:general_moment}
\end{align}
where the average is taken over $\omega$ with respect to the distribution of viewing directions. Here, we assume it to be uniform.

@@

For the third moments, we get 8 terms:
\[\begin{aligned} 
&a^3_\II[\Bell_1, \Bell_2]  \\&
 = \underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\widetilde\II[\mb i+\Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(1)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\varepsilon[\mb i+\Bell_1]\varepsilon[\mb i + \Bell_2]}_{(2)}\\ 
&+ \underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\varepsilon[\mb i + \Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(3)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\widetilde\II[\mb i + \Bell_1]\varepsilon[\mb i + \Bell_2]}_{(4)}\\
&+ \underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\widetilde\II[\mb i + \Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(5)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\varepsilon[\mb i + \Bell_1]\varepsilon[\mb i + \Bell_2]}_{(6)}\\
&+ \underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\varepsilon[\mb i + \Bell_1]\widetilde\II[\mb i + \Bell_2]}_{(7)} +
\underbrace{\frac{1}{N^2}\sum_{\mb i}\varepsilon[\mb i]\widetilde\II[\mb i + \Bell_1]\varepsilon[\mb i + \Bell_2]}_{(8)}.
\end{aligned}\]
We address these terms one by one:
\begin{itemize}
	\item  Term (1) is treated by~\eqref{eq:general_moment} for $p=3$;
	\item Term (2) is the third-order autocorrelation of  pure noise  which vanishes in the limit; 
	\item Terms (3)-(5) depend linearly on the noise and hence
	vanish in the limit;
	\item For term (6), if $\Bell_1\neq \Bell_2$ the term
	vanishes in the limit. If $\Bell_1 = \Bell_2$ then
	\[ \begin{aligned}
	\frac{1}{N^2}\sum_{\mb i}\widetilde\II[\mb i]\varepsilon[\mb i + \Bell]^2
	&= \frac{MP^2}{N^2}\cdot\frac{1}{MP^2}\sum_{t=1}^{M}\sum_{\mb i\in\mc S_t}I_{\omega_t}[\mb i]\varepsilon[\mb i + \Bell]^2 
	\\& \to \gamma\sigma^2 \langle a^1_{I_{\omega}}\rangle_{\omega}, %\gamma\mathbb{E}_{\omega, \varepsilon}[\varepsilon^2I_{\omga}]\\
	%&= \gamma\sigma^2\mathbb{E}[m_1[I_{\omega}]]\\ 
	%&= \gamma\sigma^2Lm_1[\phi],
	\end{aligned}\]
	where $\langle a^1_{I_{\omega}}\rangle_{\omega}$ is the mean of the volume. 
	\item  Terms (7) and (8) contribute $\delta$ functions similar to (6).
\end{itemize}
Thus, we conclude that
\begin{multline} \label{eq:third_order_ac_micro_volume}
a^3_\II[\Bell_1, \Bell_2] \to \gamma\langle
a^3_{I_{\omega}}[\Bell_1, \Bell_2]\rangle_{\omega} \\ +
\gamma\sigma^2\langle a^1_{I_{\omega}}\rangle_{\omega}\Big(\delta[\Bell_1 - \Bell_2] +
\delta[\Bell_1] + \delta[\Bell_2]\Big),
\end{multline}
where the second term is the bias $b_3[\Bell_1,\Bell_2]$.
Note that $\gamma\langle a^1_{I_{\omega}}\rangle_{\omega}$ is approximately the mean of the micrograph since  $a^1_\II \approx a^1_{\tilde\II} \approx \gamma\langle a^1_{I_{\omega}}\rangle_{\omega}$. Therefore, we do not need prior knowledge of $\gamma$ to effectively
debias the third-order autocorrelation.

@@

We represent our autocorrelations using Prolate Spheroidal Wave Functions (PSWFs) $\{\psi_{k,q}\}$ where $k \geq 0, q \geq 1$ are integers~\cite{slepian1964pswfIV}. As we demonstrate below, this makes it easier to account for the fact that all in-plane rotations of the micrographs are equally likely observations. This is only a concern for the second and third order autocorrelations. Below, we start with $p=2$.
The PSWFs are eigenfunctions of the truncated Fourier transform and are given in polar coordinates by\footnote{A different normalization is used in \cite{landa2017steerable}.}
\begin{align}
\psi_{k,q}(r,\varphi) & = \left\{\begin{array}{ll} \frac{1}{\sqrt{8\pi}}\alpha_{k,q}R_{k,q}(r)e^{\iota k\varphi}, & r\leq 1,\\ 0, & r>1,\end{array}\right. \label{eq:prolatesdef}
\end{align}
where the range of $k,q$ is determined by Eq.~(8) in \cite{landa2017steerable}, the ${R_{k,q}}$ are a family of real, one-dimensional functions and the ${\alpha_{k,q}}$ are  scaling factors which will be defined in the next section.
The PSWFs are orthogonal on the unit disk.

@@

We now turn out attention to the third order autocorrelation. Following the same lines, we define the coefficients:
\[\begin{aligned} \mathfrak{a}^3[k_1,q_1;k_2,q_2] &= \sum_{\Bell_1, \Bell_2} a^3_\II[\Bell_1,\Bell_2]\overline{\psi_{k_1,q_1}[\Bell_1]}\psi_{k_2,q_2}[\Bell_2]\\
&= \frac{1}{N^2}\sum_{\mb i}\II[\mb i]\left(\sum_{\Bell_1}\II[\mb i+\Bell_1]\overline{\psi_{k_1,q_1}[\Bell_1]}\right)\\&\times \left(\sum_{\Bell_2}\II[\mb i+\Bell_2]\psi_{k_2,q_2}[\Bell_2]\right)\\
&= \frac{1}{N^2}\sum_{\mb i}\II[\mb i] a_{k_1,q_1}[\mb i]\overline{a_{k_2,q_2}[\mb i]},
\end{aligned}\]
where the patch expansion coefficients $a_{k,q}$ are as defined in~\eqref{eq:patchcoefficients}.
The coefficients $\mathfrak{a}^3[k_1,q_1;k_2,q_2]$ are related to the third-order autocorrelation via the approximate identity:
\[ a^3_\II[\Bell_1, \Bell_2] \approx \sum_{\substack{k_1,q_1\\ k_2,q_2}}\mathfrak{a}^3[k_1,q_1;k_2,q_2]\psi_{k_1,q_1}[\Bell_1]\overline{\psi_{k_2,q_2}[\Bell_2]}.\]
Averaging over all rotations of $\II$ and its reflection, we obtain
\begin{align} \label{eq:third_order_ac_pswf}
&\mathfrak{a}^3[k_1,q_1;k_2,q_2] = \frac{1}{N^2}\sum_{\mb i}\II[\mb i]\\&\times \left(\frac{1}{4\pi}\int_0^{2\pi}\left(a_{k_1,q_1}[\mb i]\overline{a_{k_2,q_2}[\mb i]} + \overline{a_{k_1,q_1}[\mb i]}a_{k_2,q_2}[\mb i]\right)e^{-\iota(k_1-k_2)\alpha}d\alpha\right) \nonumber\\
&= \delta[k_1 - k_2]\frac{1}{N^2}\sum_{\mb i}\II[\mb i]\Re\{a_{k_1,q_1}[\mb i]\overline{a_{k_2,q_2}[\mb i]}\}. \nonumber 
\end{align}
Thus, similarly to the second-order autocorrelation, averaging over all in-plane rotations reveals that the 4-D third-order autocorrelation truly only carries information along three dimensions.
